Multi-State Temporal Occupancy Modeling

### 2020-07-29 Model Statement/MSTOM Statement.tex
updated model statement file

\begin{document}
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{0pt}
\setlength{\abovedisplayshortskip}{0pt}
\setlength{\belowdisplayshortskip}{0pt}
\maketitle

@@

We could also model detection parameters by site (\textbf{x}) or survey specific (\textbf{V}) covarariates, such that 
\begin{center}
\begin{align*}
\boldsymbol{q}_{ij} &= \begin{bmatrix} 1 & 0 & 0 & 0 \\ 
									1 &  q^{2,2}_{ij} & 0 & 0 \\ 
									1 & 0 & q^{3,3}_{ij} & 0\\
  								       1 & q^{4,2}_{ij} & q^{4,3}_{ij} & q^{4,4}_{ij}
  								      \end{bmatrix}\\
q^{2,2}_{ij} &=e^{\beta^{\text{Day}}_{0} +\beta^{\text{Day}}_{1}\times x_{i}+\beta^{\text{Day}}_{2}\times v_{ij}}\\
q^{3,3}_{ij} &=e^{\beta^{\text{Night}}_{0} +\beta^{\text{Night}}_{1}\times x_{i}+\beta^{\text{Day}}_{2}\times v_{ij}}\\
q^{4,2}_{ij} &=e^{\beta^{\text{ND2}}_{0} +\beta^{\text{ND2}}_{1}\times x_{i}+\beta^{\text{Day}}_{2}\times v_{ij}}\\
q^{4,3}_{ij} &=e^{\beta^{\text{ND3}}_{0} +\beta^{\text{ND3}}_{1}\times x_{i}+\beta^{\text{Day}}_{2}\times v_{ij}}\\
q^{4,4}_{ij} &=e^{\beta^{\text{ND4}}_{0} +\beta^{\text{ND4}}_{1}\times x_{i}+\beta^{\text{Day}}_{2}\times v_{ij}}
\end{align*}
\end{center}
and the probability of detection parameters can be derived as, 
\begin{center}
\begin{align*}
p^{\text{Day}}_{ij} &=\frac{q^{2,2}_{ij}}{1+q^{2,2}_{ij}+0+0}\\
p^{\text{Night}}_{ij} &=\frac{q^{3,3}_{ij}}{1+0+q^{3,3}_{ij}+0}\\
p^{\text{ND1}}_{ij} &=\frac{1}{1+q^{4,2}_{ij}+q^{4,3}_{ij}+q^{4,4}_{ij}}\\
p^{\text{ND2}}_{ij} &=\frac{q^{4,2}_{ij}}{1+q^{4,2}_{ij}+q^{4,3}_{ij}+q^{4,4}_{ij}}\\
p^{\text{ND3}}_{ij} &=\frac{q^{4,3}_{ij}}{1+q^{4,2}_{ij}+q^{4,3}_{ij}+q^{4,4}_{ij}}\\
p^{\text{ND4}}_{ij} &=\frac{q^{4,4}_{ij}}{1+q^{4,2}_{ij}+q^{4,3}_{ij}+q^{4,4}_{ij}}.
\end{align*}
\end{center}

Diffuse priors for all $\beta$ parameters could be specifed as Logistic(0,1).

@@

Similarly, we can simplify the detection matrix using the probability of detection during the day  ($p^{\text{Day.M}}$) and night ($p^{\text{Night.M}}$) as,
\begin{equation}
\boldsymbol{p}_{j} = \begin{bmatrix} 1 & 0 & 0 & 0 \\ 
									1-p^{\text{Day.M}} & p^{\text{Day.M}} & 0 & 0 \\ 
									1-p^{\text{Night.M}} & 0 & p^{\text{Night.M}} & 0\\
  								      (1-p^{\text{Day.M}})(1-p^{\text{Night.M}}) & p^{\text{Day.M}}(1-p^{\text{Night.M}}) & (1-p^{\text{Day.M}})p^{\text{Night.M}} & p^{\text{Day.M}}p^{\text{Night.M}}
  								      \end{bmatrix}
\end{equation}
, where $p^{1,4}$ can also be specified as equal to $1- p^{\text{Day.M}}(1-p^{\text{Night.M}}) - (1-p^{\text{Day.M}})p^{\text{Night.M}} - p^{\text{Day.M}}p^{\text{Night.M}}$. Unless the species of interest is completely cathemeral and uses the day and night times in equal proportion, it is unlikely that $p^{\text{Day.M}}$ and $p^{\text{Night.M}}$ should be equivalent. We could use the diffuse prior distributions of 
\begin{center}
\begin{align*}
\psi^{\text{Day.M}} &\sim \text{Beta}(1,1)\\
\psi^{\text{Night.M}} &\sim \text{Beta}(1,1)\\
p^{\text{Day.M}} &\sim \text{Beta}(1,1)\\
p^{\text{Night.M}} &\sim \text{Beta}(1,1).\\
\end{align*}
\end{center}
This model is coded for JAGS in the file `jags.multistate.occ.reduced.R'.

>>>>

\begin{document}
\setlength{\abovedisplayskip}{-2pt}
\setlength{\belowdisplayskip}{10pt}
\setlength{\abovedisplayshortskip}{0pt}
\setlength{\belowdisplayshortskip}{10pt}
\maketitle

@@

We could also model detection parameters by site (\textbf{x}) or survey specific (\textbf{V}) covarariates, such that 
\begin{center}
\begin{align*}
\boldsymbol{q}_{ij} &= \begin{bmatrix} 1 & 0 & 0 & 0 \\ 
									1 &  q^{2,2}_{ij} & 0 & 0 \\ 
									1 & 0 & q^{3,3}_{ij} & 0\\
  								       1 & q^{4,2}_{ij} & q^{4,3}_{ij} & q^{4,4}_{ij}
  								      \end{bmatrix}\\
q^{2,2}_{ij} &=e^{\alpha^{\text{Day}}_{0} +\alpha^{\text{Day}}_{1}\times x_{i}+\alpha^{\text{Day}}_{2}\times v_{ij}}\\
q^{3,3}_{ij} &=e^{\alpha^{\text{Night}}_{0} +\alpha^{\text{Night}}_{1}\times x_{i}+\alpha^{\text{Day}}_{2}\times v_{ij}}\\
q^{4,2}_{ij} &=e^{\alpha^{\text{ND2}}_{0} +\alpha^{\text{ND2}}_{1}\times x_{i}+\alpha^{\text{Day}}_{2}\times v_{ij}}\\
q^{4,3}_{ij} &=e^{\alpha^{\text{ND3}}_{0} +\alpha^{\text{ND3}}_{1}\times x_{i}+\alpha^{\text{Day}}_{2}\times v_{ij}}\\
q^{4,4}_{ij} &=e^{\alpha^{\text{ND4}}_{0} +\alpha^{\text{ND4}}_{1}\times x_{i}+\alpha^{\text{Day}}_{2}\times v_{ij}}
\end{align*}
\end{center}
and the probability of detection parameters can be derived as, 
\begin{center}
\begin{align*}
p^{\text{Day}}_{ij} &=\frac{q^{2,2}_{ij}}{1+q^{2,2}_{ij}+0+0}\\
p^{\text{Night}}_{ij} &=\frac{q^{3,3}_{ij}}{1+0+q^{3,3}_{ij}+0}\\
p^{\text{ND1}}_{ij} &=\frac{1}{1+q^{4,2}_{ij}+q^{4,3}_{ij}+q^{4,4}_{ij}}\\
p^{\text{ND2}}_{ij} &=\frac{q^{4,2}_{ij}}{1+q^{4,2}_{ij}+q^{4,3}_{ij}+q^{4,4}_{ij}}\\
p^{\text{ND3}}_{ij} &=\frac{q^{4,3}_{ij}}{1+q^{4,2}_{ij}+q^{4,3}_{ij}+q^{4,4}_{ij}}\\
p^{\text{ND4}}_{ij} &=\frac{q^{4,4}_{ij}}{1+q^{4,2}_{ij}+q^{4,3}_{ij}+q^{4,4}_{ij}}.
\end{align*}
\end{center}

Diffuse priors for all $\alpha$ parameters could be specifed as Logistic(0,1).

@@

Similarly, we can simplify the detection matrix using the probability of detection during the day  ($p^{\text{Day.M}}$) and night ($p^{\text{Night.M}}$) as,
\vspace{10pt}
\begin{equation}
\boldsymbol{p}_{j} = \begin{bmatrix} 1 & 0 & 0 & 0 \\ 
									1-p^{\text{Day.M}} & p^{\text{Day.M}} & 0 & 0 \\ 
									1-p^{\text{Night.M}} & 0 & p^{\text{Night.M}} & 0\\
  								      (1-p^{\text{Day.M}})(1-p^{\text{Night.M}}) & p^{\text{Day.M}}(1-p^{\text{Night.M}}) & (1-p^{\text{Day.M}})p^{\text{Night.M}} & p^{\text{Day.M}}p^{\text{Night.M}}
  								      \end{bmatrix}
\end{equation}
, where $p^{1,4}$ can also be specified as equal to $1- p^{\text{Day.M}}(1-p^{\text{Night.M}}) - (1-p^{\text{Day.M}})p^{\text{Night.M}} - p^{\text{Day.M}}p^{\text{Night.M}}$. Unless the species of interest is completely cathemeral and uses the day and night times in equal proportion, it is unlikely that $p^{\text{Day.M}}$ and $p^{\text{Night.M}}$ should be equivalent. We could use the diffuse prior distributions of 
\begin{center}
\begin{align*}
\psi^{\text{Day.M}} &\sim \text{Beta}(1,1)\\
\psi^{\text{Night.M}} &\sim \text{Beta}(1,1)\\
p^{\text{Day.M}} &\sim \text{Beta}(1,1)\\
p^{\text{Night.M}} &\sim \text{Beta}(1,1).\\
\end{align*}
\end{center}
This model is coded for JAGS in the file `jags.multistate.occ.reduced.R'.
